Kids SUMs Summary

### 2020-02-06 pm_summary.Rmd
SUMs cleaning updates

---
title: "Kids PM Summary"
author: "Ethan Walker"
date: "Started 22 Nov 2019, Updated 27 Jan 2020"
output: pdf_document
---

@@

# Summary stats
```{r}
pm_summary <- pm_clean %>% 
  # group by: area, treatment, dtid, day_of_week, Sampling_day
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  filter(pm < 500) %>% 
  group_by(area) %>% 
  summarize("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "25%" = quantile(pm, 0.25),
            "Median PM" = median(pm, na.rm = TRUE),
            "75%" = quantile(pm, 0.75),
            "Max PM" = max(pm, na.rm = TRUE),
            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
            "Mean % Daily Obs" = mean(percent_daily_obs_pm, na.rm = TRUE)) %>% 
  select(-area) %>% 
  t()

kable(pm_summary, col.names = c("Alaska", "Navajo", "Montana"), 
      digits = 1, align = "c")
```

>>>>

---
title: "Kids PM Summary"
author: "Ethan Walker"
date: "Started 22 Nov 2019, Updated 2 Feb 2020"
output: pdf_document
---

@@

# Summary stats
```{r}
pm_summary <- pm_clean %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(pm < 500) %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  group_by(area) %>% 
  summarize("n" = n(),
            "Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "25%" = quantile(pm, 0.25, na.rm = TRUE),
            "Median PM" = median(pm, na.rm = TRUE),
            "75%" = quantile(pm, 0.75, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE),
            "Mean Sample Days" = mean(pm_sample_interval, na.rm = TRUE),
            "SD Sample Days" = sd(pm_sample_interval, na.rm = TRUE),
            "Mean Sample Obs" = mean(pm_total_observations, na.rm = TRUE),
            "Mean % Daily Obs" = mean(percent_daily_obs_pm, na.rm = TRUE)) %>% 
  #select(-treatment) %>% 
  t()

kable(pm_summary, 
      digits = 1, align = "c")

pm_by_day <- pm_clean %>% 
  mutate(sampling_month = month(pm_datetime_new, abbr = TRUE)) %>% 
  mutate(day_of_week = factor(day_of_week,
                                 levels = c("Monday", "Tuesday", "Wednesday", "Thursday",
                                            "Friday", "Saturday", "Sunday"),
                                 labels = c("Mon", "Tues", "Wed", "Thur", "Fri", "Sat", "Sun"))) %>% 
  filter(winter_id == 1 & treatment == "Placebo") %>% 
  #filter(pm < 500) %>% 
  mutate(pm = if_else(pm > 500, 500, pm)) %>% 
  group_by(area, day_of_week) %>% 
  summarize("Mean PM" = mean(pm, na.rm = TRUE), 
            "SD PM" = sd(pm, na.rm = TRUE),
            "Min PM" = min(pm, na.rm = TRUE), 
            "25%" = quantile(pm, 0.25, na.rm = TRUE),
            "Median PM" = median(pm, na.rm = TRUE),
            "75%" = quantile(pm, 0.75, na.rm = TRUE),
            "Max PM" = max(pm, na.rm = TRUE))
pm_by_day
```

### 2020-02-06 sums_data_cleaning.Rmd
SUMs cleaning updates

---
title: "KidsAIR initial SUMs data work"
author: "Ethan Walker"
date: "Started 23 Dec 2019, Updated 30 Jan 2020"
output: pdf_document
---

@@

# write_rds(sums_clean, "Output/sums_clean.rds")
```

>>>>

---
title: "KidsAIR initial SUMs data work"
author: "Ethan Walker"
date: "Started 23 Dec 2019, Updated 6 Feb 2020"
output: pdf_document
---

@@

# write_rds(sums_clean, "Output/sums_clean.rds")

########## New cleaning on 6 Feb 2020 ##########

sums_clean <- read_rds("Output/sums_clean.rds")

# This cleaning is a bit of a desperation attempt to try and salvage most of the
# SUMs data. Because of the way data was collected and files were saved, it
# became very difficult ot piece files together for each home. For some of the homes,
# data is fine - particularly if they used LogTags for the entire Winter and there
# is one continuous file from one instrument for the whole Winter. However, if an
# IButton was used (which held only around 1 month of data), or if a LogTag was
# switched out during sampling, piecing the various files together becomes a nightmare.
# All of the files had heads and tails of pre- or post-sampling data that needed
# to be cut off, but there wasn't a good indication of when the sampling exactly
# stopped and started when an instrument was switched out in the field. There was
# typically overlap in the files when an instrument was switched out: a tail from
# the old instrument before it was turned off, and a head from the new instrument
# from the time it was turned on, but prior to sampling. During these overlap
# periods in the data, there was no good indication of what instrument was actually
# sampling within the home. I tried to write some rules/algorithms to pull out what
# I thought was the correct data, but there inconsistencies were inconsistent. 
# Nothing I did made the data seem trustworthy for most of the homes, so here I have
# resorted to cutting out those overlapping periods in the data. I am filtering out
# around 161,000 observations (of 3.5 million, so < 5% of the data), but I can't come
# up with any other way to have cleaned data that I actually trust, short of 
# pasting each individual file together by hand. Even then, there are formatting
# issues that would take days to resolve. So, this is the best way I have found
# to get what I think is reliable SUMs data: filtering out the observations
# with overlapping datapoints by pulling out instances when the time between
# sampling points is not consistent.

sums_new <- sums_clean %>% 
  group_by(area, home) %>% 
  mutate(time_diff = lead(datetime_sums) - datetime_sums,
         time_diff_indicator = if_else(time_diff != rollmean(time_diff, 4), 1, 0)) %>% 
  filter(time_diff_indicator != 1) %>%
  ungroup() %>% 
  arrange(area, home, winter_id, datetime_sums)

# write_rds(sums_new, "Output/sums_clean.rds")
```

### 2020-02-06 sums_summary.Rmd
SUMs cleaning updates

---
title: "Kids SUMs Summary"
author: "Ethan Walker"
date: "Started 9 Jan 2020, Updated 30 Jan 2020"
output: pdf_document
---

@@

thresholds_function(sums_filtered_date, 30, 1.5)
thresholds_function(sums_filtered_date, 40, 1.5)
thresholds_function(sums_filtered_date, 50, 1.5)
```

@@

thresholds <- sums_filtered_date %>% 
  group_by(area, home, winter_id, date_sums) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= 40, 1, 0)) %>% 
  summarize(percent_above = sum(threshold_temp)/n()*100) %>% 
  group_by(area) %>% 
  summarize("% time/day > threshold" = round(mean(percent_above), digits = 2))
thresholds

@@

thresholds_percent(sums_filtered_date, 25)
thresholds_percent(sums_filtered_date, 30)
thresholds_percent(sums_filtered_date, 35)
thresholds_percent(sums_filtered_date, 40)
thresholds_percent(sums_filtered_date, 45)
thresholds_percent(sums_filtered_date, 50)
thresholds_percent(sums_filtered_date, 55)
```

>>>>

---
title: "Kids SUMs Summary"
author: "Ethan Walker"
date: "Started 9 Jan 2020, Updated 7 Feb 2020"
output: pdf_document
---

@@

thresholds_function(sums_filtered_date, 25, 1.5)
thresholds_function(sums_filtered_date, 27, 1.5)
thresholds_function(sums_filtered_date, 30, 1.5)
```

@@

thresholds <- data %>% 
  group_by(area, home, winter_id, date_sums) %>%
  mutate(threshold_temp = if_else(temp_c_sums >= temp, 1, 0)) %>% 
  summarize(percent_above = sum(threshold_temp)/n()*100) %>% 
  group_by(area) %>% 
  summarize("% time/day > threshold" = round(mean(percent_above), digits = 2))
thresholds

@@

thresholds_percent(sums_filtered_date, 25)
thresholds_percent(sums_filtered_date, 27)
thresholds_percent(sums_filtered_date, 30)
```
